R Notebook

### 2023-11-16 Tensor MSSA.rmd
Add files via upload

# rank of the time series
r <- 2
# generating series with ranodm parameters
s <- map(1:series.number, function(x) runif(1, 1, 1) * sin(1:N * pi / 3))
# generating series by hand
# s <- map(1:series.number, function(x) exp(1:N))
s.hankel <- lapply(s, hankel, L)
mat <- reduce(s.hankel, cbind)
u.svd <- svd(mat)
# signal subspace
true.subspace <- u.svd$u[, 1:r, drop = FALSE]
```

@@

  s.tens <- fold(mat.n, 1, 2:3, modes = c(L, N - L + 1, series.number))
  capture.output({ h <- rTensor::hosvd(s.tens) })
  tens <- ttl(h$Z[1:r, 1:r, 1:r, drop = FALSE], list(h$U[[1]][, 1:r, drop = FALSE], h$U[[2]][, 1:r, drop = FALSE], h$U[[3]][, 1:r, drop = FALSE]), 1:3)
  # tens <- as.tensor(h$Z@data[1, 1, 1] * (h$U[[1]][, 1] %o% h$U[[2]][, 1] %o% h$U[[3]][, 1]) + h$Z@data[2, 2, 2] * (h$U[[1]][, 2] %o% h$U[[2]][, 2] %o% h$U[[3]][, 2]))
  est.subspace.tens <- unfold(tens, 1, 2:3)@data
  tens.rec <- apply(tens@data, 3, hankel, simplify = FALSE)

@@

set.seed(1)
N <- 71
sigma <- 5
Ls <- c(12, 24, 36, 48, 60)
signal1 <- 30 * cos(2 * pi * 1:N / 12)
# signal2 <- 20 * cos(2 * pi * 1:N / 8 + pi / 4)
signal2 <- 20 * cos(2 * pi * 1:N / 12 + pi / 4)
signal <- cbind(signal1, signal2)
R <- 100
r <- 2
mssa.errors <- function(Ls) {
  f1 <- signal1 + rnorm(N, sd = sigma)
  f2 <- signal2 + rnorm(N, sd = sigma)
  f <- list(f1, f2)
  err.rec <- list(mat = numeric(length(Ls)), tens = numeric(length(Ls)))
  names(err.rec$mat) <- Ls; names(err.rec$tens) <- Ls
  for (l in seq_along(Ls)) {
    L <- Ls[l]
    s <- ssa(f, L = L, kind = "mssa")
    rec <- reduce(reconstruct(s, groups = list(1:r))[[1]], cbind)
    err.rec$mat[l] <- mean((rec - signal)^2)

    f.hankel <- lapply(f, hankel, L)
    mat.f <- reduce(f.hankel, cbind)
    s.tens <- fold(mat.f, 1, 2:3, modes = c(L, N - L + 1, 2))
    capture.output({ h <- rTensor::hosvd(s.tens) })
    tens <- ttl(h$Z[1:r, 1:r, 1:min(r, 2), drop = FALSE], list(h$U[[1]][, 1:r, drop = FALSE], h$U[[2]][, 1:r, drop = FALSE], h$U[[3]][, 1:min(r, 2), drop = FALSE]), 1:3)
    rec.tens <- reduce(apply(tens@data, 3, hankel, simplify = FALSE), cbind)
    err.rec$tens[l] <- mean((rec.tens - signal)^2)
  }
  err.rec
}
mres <- replicate(R, mssa.errors(Ls))
```

Среднее MSE для матричного и тензорного подходов для различных длин окна L
```{r}
rowMeans(simplify2array(mres[1,]))
rowMeans(simplify2array(mres[2,]))
```

>>>>

# rank of the time series
r <- 2
# generating series with ranodm parameters
s <- map(1:series.number, function(x) runif(1, 1, 4) * sin(1:N * pi / 3))
# generating series by hand
# s <- map(1:series.number, function(x) exp(1:N))
s.hankel <- lapply(s, hankel, L)
mat <- reduce(s.hankel, cbind)
u.svd <- svd(mat)
# signal subspace
true.subspace <- u.svd$u[, 1:r, drop = FALSE]
```

@@

  s.tens <- fold(mat.n, 1, 2:3, modes = c(L, N - L + 1, series.number))
  capture.output({ h <- rTensor::hosvd(s.tens) })
  tens <- ttl(h$Z[1:r, 1:r, 1, drop = FALSE], list(h$U[[1]][, 1:r, drop = FALSE], h$U[[2]][, 1:r, drop = FALSE], h$U[[3]][, 1, drop = FALSE]), 1:3)
  # tens <- as.tensor(h$Z@data[1, 1, 1] * (h$U[[1]][, 1] %o% h$U[[2]][, 1] %o% h$U[[3]][, 1]) + h$Z@data[2, 2, 2] * (h$U[[1]][, 2] %o% h$U[[2]][, 2] %o% h$U[[3]][, 2]))
  est.subspace.tens <- unfold(tens, 1, 2:3)@data
  tens.rec <- apply(tens@data, 3, hankel, simplify = FALSE)

@@

set.seed(1)
N <- 71
sigma <- 5
series.numbers <- c(2, 8, 14, 20)
# series.numbers <- 2
Ls <- c(12, 24, 36, 48, 60)
r <- 2
# signal1 <- 30 * cos(2 * pi * 1:N / 12)
# signal2 <- 20 * cos(2 * pi * 1:N / 12 + pi / 4)
# signal2 <- 20 * cos(2 * pi * 1:N / 8 + pi / 4)
# signal3 <- 40 * cos(2 * pi * 1:N / 12 + pi / 20)
mres <- list()
for (series.number in series.numbers) {
  signal_l <- replicate(series.number, runif(1, 20, 35) * cos(2 * pi * 1:N + pi / runif(1, 2, 20)), simplify = FALSE)
  signal <- reduce(signal_l, cbind)
  # signal <- cbind(signal1, signal2, signal3)
  R <- 100

  mssa.errors <- function(Ls) {
    # f1 <- signal1 + rnorm(N, sd = sigma)
    # f2 <- signal2 + rnorm(N, sd = sigma)
    # f3 <- signal3 + rnorm(N, sd = sigma)
    # f <- list(f1, f2, f3)
    f <- lapply(signal_l, function(x) x + rnorm(N, sd = sigma))
    err.rec <- list(mat = numeric(length(Ls)), tens = numeric(length(Ls)))
    names(err.rec$mat) <- Ls; names(err.rec$tens) <- Ls
    for (l in seq_along(Ls)) {
      L <- Ls[l]
      s <- ssa(f, L = L, kind = "mssa")
      rec <- reduce(reconstruct(s, groups = list(1:r))[[1]], cbind)
      err.rec$mat[l] <- mean((rec - signal)^2)

      f.hankel <- lapply(f, hankel, L)
      mat.f <- reduce(f.hankel, cbind)
      s.tens <- fold(mat.f, 1, 2:3, modes = c(L, N - L + 1, series.number))
      capture.output({ h <- rTensor::hosvd(s.tens) })
      tens <- ttl(h$Z[1:r, 1:r, 1:series.number, drop = FALSE], list(h$U[[1]][, 1:r, drop = FALSE], h$U[[2]][, 1:r, drop = FALSE], h$U[[3]][, 1:series.number, drop = FALSE]), 1:3)
      rec.tens <- reduce(apply(tens@data, 3, hankel, simplify = FALSE), cbind)
      err.rec$tens[l] <- mean((rec.tens - signal)^2)
    }
    err.rec
  }

  # mres <- replicate(R, mssa.errors(Ls))
  mres[[toString(series.number)]] <- replicate(R, mssa.errors(Ls))
}
```

Среднее MSE для матричного и тензорного подходов для различных длин окна L
```{r}
for (i in series.numbers) {
  print(i)
  print(rowMeans(simplify2array(mres[[toString(i)]][1,])))
  print(rowMeans(simplify2array(mres[[toString(i)]][2,])))
}
```

```{r}
rowMeans(simplify2array(mres[1,]))
rowMeans(simplify2array(mres[2,]))
```
